---
title: "Raster geoms"
author: "Viktor Petukhov"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Raster points
Sometimes you need to publish a figure in a fector format. But your figure contains thousands of points:
```{r, fig.width=4, fig.height=4}
library(ggplot2)

points_num <- 500000
df <- data.frame(x=rnorm(points_num), y=rnorm(points_num), c=as.factor(1:points_num %% 2))
gg <- ggplot(df, aes(x=x, y=y, color=c)) + scale_color_discrete(guide=F)

(gg_vec <- gg + geom_point(size=0.5))
```

In this case the reasonable solution would be to rasterize the plot. But in this case all text becomes raster as well.
Raster layers were developed to prevent such situation:
```{r, fig.width=4, fig.height=4}
(gg_rast <- gg + geom_point_rast(size=0.5))
```

Plots looks the same, but difference can be seen when they are exported to pdfs. Unfortunately, the price is longer rendering time.
```{r}
PrintFileSize <- function(gg, name) {
  pdf('tmp.pdf', width = 4, height = 4)
  print(gg)
  invisible(dev.off())
  cat(name, ': ', file.info('tmp.pdf')$size / 1024, ' Kb.\n', sep = '')
  unlink('tmp.pdf')
}

PrintFileSize(gg_rast, 'Raster')
PrintFileSize(gg_vec, 'Vector')
```

As expected, difference becomes larger with growth of number of points:
```{r}
points_num <- 1000000
df <- data.frame(x=rnorm(points_num), y=rnorm(points_num), c=as.factor(1:points_num %% 2))
gg <- ggplot(df, aes(x=x, y=y, color=c)) + scale_color_discrete(guide=F)

gg_vec <- gg + geom_point(size=0.5)
gg_rast <- gg + geom_point_rast(size=0.5)

PrintFileSize(gg_rast, 'Raster')
PrintFileSize(gg_vec, 'Vector')
```

## Troubleshooting
In current version, legends can raster disturbe plots:
```{r, fig.width=2, fig.height=4}
points_num <- 10000
df <- data.frame(x=rnorm(points_num), y=rnorm(points_num), c=as.factor(1:points_num %% 2))
ggplot(df, aes(x=x, y=y, color=c)) + geom_point_rast(size=0.5)
```

To restore sides ratio, parameters width and height can be used:
```{r, fig.width=2, fig.height=4}
points_num <- 10000
df <- data.frame(x=rnorm(points_num), y=rnorm(points_num), c=as.factor(1:points_num %% 2))
ggplot(df, aes(x=x, y=y, color=c)) + geom_point_rast(size=0.5, width = 1)
```

# Boxplot outliers
Another type of plots with potentially large number of small objects is geom_boxplot:
```{r, fig.width=5, fig.height=4}
points_num <- 1000000
df <- data.frame(x=as.factor(1:points_num %% 2), y=log(abs(rcauchy(points_num))))
gg <- ggplot(df, aes(x=x, y=y)) + scale_color_discrete(guide=F)

gg + geom_boxplot()
```

With large number of objects, outlier points become noninformative. It's better to jitter them:
```{r, fig.width=4, fig.height=3}
gg_vec <- gg + geom_boxplot_jitter(outlier.size=0.1, outlier.jitter.width = 0.3, outlier.alpha=0.5)
gg_vec
```

And this geom can be rasterized as well:
```{r, fig.width=4, fig.height=3}
gg_rast <- gg + geom_boxplot_jitter(outlier.size=0.1, outlier.jitter.width = 0.3, outlier.alpha=0.5, raster=T, raster.dpi = 200)
gg_rast
```


```{r}
PrintFileSize(gg_rast, 'Raster')
PrintFileSize(gg_vec, 'Vector')
```

