---
title: "Raster geoms"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{"Raster geoms"}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Raster points
### Points
Sometimes you need to publish a figure in a vector format:
```{r, fig.width=4, fig.height=4}
library(ggplot2)
library(ggrastr)

points_num <- 50  
df <- data.frame(x=rnorm(points_num), y=rnorm(points_num), c=as.factor(1:points_num %% 2))
gg <- ggplot(df, aes(x=x, y=y, color=c)) + scale_color_discrete(guide=F)

(gg_vec <- gg + geom_point(size=0.5))
```

But in other cases, your figure contains thousands of points, e.g. try `points_num <- 500000` in the example above, and you will notice the performance issues. 
In this case, the reasonable solution would be to rasterize the plot. But the problem is that all text becomes raster as well.
Raster layers were developed to prevent such a situation:
```{r, fig.width=4, fig.height=4}
(gg_rast <- gg + geom_point_rast(size=0.5))
```

The plots look the same, but the difference can be seen when they are exported to pdfs. Unfortunately, the price is longer rendering time.
```{r}
PrintFileSize <- function(gg, name) {
  invisible(ggsave('tmp.pdf', gg, width = 4, height = 4))
  cat(name, ': ', file.info('tmp.pdf')$size / 1024, ' Kb.\n', sep = '')
  unlink('tmp.pdf')
}

PrintFileSize(gg_rast, 'Raster')
PrintFileSize(gg_vec, 'Vector')
```

As expected, the difference becomes larger with growth of number of points:
```{r}
points_num <- 1000000
df <- data.frame(x=rnorm(points_num), y=rnorm(points_num), c=as.factor(1:points_num %% 2))
gg <- ggplot(df, aes(x=x, y=y, color=c)) + scale_color_discrete(guide=F)

gg_vec <- gg + geom_point(size=0.5)
gg_rast <- gg + geom_point_rast(size=0.5)

PrintFileSize(gg_rast, 'Raster')
PrintFileSize(gg_vec, 'Vector')
```

### Tile
Heatmaps also don't work well with vector formats:
```{r}
coords <- expand.grid(1:100, 1:100)
coords$Value <- 1 / apply(as.matrix(coords), 1, function(x) sum((x - c(50, 50))^2)^0.01)
ggplot(coords) + geom_tile(aes(x=Var1, y=Var2, fill=Value))
ggplot(coords) + geom_tile_rast(aes(x=Var1, y=Var2, fill=Value))
```


### Boxplot outliers
Another type of plots with a potentially large number of small objects is geom_boxplot:
```{r, fig.width=5, fig.height=4}
points_num <- 100
df <- data.frame(x=as.factor(1:points_num %% 2), y=log(abs(rcauchy(points_num))))
gg <- ggplot(df, aes(x=x, y=y)) + scale_color_discrete(guide=F)

gg + geom_boxplot()
```

Try the above example with `points_num <- 100000` or a larger integer---it will take quite long to render, far longer than allowed via CRAN. 

With large number of objects, outlier points become noninformative. It's better to jitter them:
```{r, fig.width=4, fig.height=3}
gg_vec <- gg + geom_boxplot_jitter(outlier.size=0.1, outlier.jitter.width = 0.3, outlier.alpha=0.5)
gg_vec
```

And this geom can be rasterized as well:
```{r, fig.width=4, fig.height=3}
gg_rast <- gg + geom_boxplot_jitter(outlier.size=0.1, outlier.jitter.width = 0.3, outlier.alpha=0.5, raster=T, raster.dpi = 200)
gg_rast
```


```{r}
PrintFileSize(gg_rast, 'Raster')
PrintFileSize(gg_vec, 'Vector')
```

## Troubleshooting
In the current version, legends can disturb raster plots:
```{r, fig.width=2, fig.height=4}
points_num <- 10
df <- data.frame(x=rnorm(points_num), y=rnorm(points_num), c=as.factor(1:points_num %% 2))
ggplot(df, aes(x=x, y=y, color=c)) + geom_point_rast(size=0.5)
```

To restore the side ratio, teh parameters width and height can be used:
```{r, fig.width=2, fig.height=4}
points_num <- 10
df <- data.frame(x=rnorm(points_num), y=rnorm(points_num), c=as.factor(1:points_num %% 2))
ggplot(df, aes(x=x, y=y, color=c)) + geom_point_rast(size=0.5, raster.width = 1)
```